---
title: "R package DiversityOccu: Alpha diversity modeled from Occupancy"
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R package DiversityOccu: Alpha diversity modeled from Occupancy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#Modeling diversity from Occupancy.

Several authors have discussed the possibility of managing based on species 
desitions or community. This package develops tools, that make the use of both 
aproaches at the same time using occupancy modelling to predict species diversity
indices, model diversity, and take both individual species modelling and species
diviersity index modelling, in order to take desitions based on both a community
and individual species background.

##Theory

Since its inception in 2002 by MacKenzie et al. 

##How to use it

###Data formating

In order to use the DiversityOccupancy package you need at least three files:

####Detection history

###Fiting the models and predicting diversity

###Selecting the best model to predict diversity

###Visualising the response of diversity to the selected Variables

###Predicting to a new dataset 

###Spatial data

#References

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
