---
title: "R package DiversityOccu: Alpha diversity modeled from Occupancy"
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R package DiversityOccu: Alpha diversity modeled from Occupancy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#Modeling diversity from Occupancy.

Several authors have discussed the possibility of managing based on species 
desitions or community. This package develops tools, that make the use of both 
aproaches at the same time using occupancy modelling to predict species diversity
indices, model diversity, and take both individual species modelling and species
diviersity index modelling, in order to take desitions based on both a community
and individual species background.

##Theory

Since its inception in 2002 by MacKenzie et al. 

##How to use it

###Data formating
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(DiversityOccupancy)
library(knitr)
```
In order to use the DiversityOccupancy package you need at least three files:

####Detection history of multiple species

A data frame consisting on the detection history of at least one species, as an 
example *DiversityOccupancy* has the dataset IslandBirds, a dataset containing 
the detections history of 13 species in the Pohnpei Island for 4 consecutive days 
(Columns) in 120 different sites (Rows). For every time the species was detected
you have a 1 and for every time the species was not, you have a 0.

```{r, echo=TRUE}
data("IslandBirds")
head(IslandBirds)
```

####Site covariates

A data frame consisting on measurements taken on each site, this covariates are 
used to model occupnacy or abundance, and are pur variables of interest, this are
supposed to be variables that are stable within the scope of the length of the 
study. In *DiversityOccupancy* there is an example concordant with the IslandBids
dataset called siteCov

```{r, echo=TRUE}
data("siteCov")
head(siteCov)
```

####Daily covariates

A list of data frames, where each data frame is a daily measurement of a variable
thought to affect the probability to detect the species. It is important that each 
element (data frame) of the list, to have a name so that it can be called to fit 
the occupancy model. This variables are used to model the probability of detection.

*DiversityOccupancy* has a dataset called Daily_Cov which ilustrates how the Daily
covariates have to be structured

```{r, echo=TRUE}
summary(Daily_Cov)
names(Daily_Cov)
head(Daily_Cov[[1]])
```

###Fiting the models and predicting diversity

####Modeling alpha diversity

In order to model alpha diversity, we use the function *diversityoccu*. This function fits the latent abundance mixture model described in Royle and Nichols (2003), to calculate the abundance of every species in each site, the using that abundance it calculates the alpha diversity index for each site based on that abundance. If 
dredge = TRUE, the function will automatically select the best model by fitting
all posible models and selecting the best one using AICc. The calculated diversity 
index can be chosen to be either shannon's (default), simpson, or inverse simpson.
The formula for the full model has to be stipullated following the format ~ obscov 
~ sitcov, the first arguments will be used to calculate probability of detection 
and the second part occupancy. The number of species is stated in the parameter
spp.

The following example shows how to model Bird Diversity using part of the variables
of the siteCov and Daily_Cov datasets, predicting shannon's diversity index and 
selecting model by AICc.

```{r, echo=TRUE, cache=TRUE, warning=FALSE, message=FALSE}
BirdDiversity<-diversityoccu(pres = IslandBirds, sitecov = siteCov, obscov = Daily_Cov, spp = 13, dredge = FALSE, form = ~ Wind + Rain + Noise ~ Elev + RPR + AgroFo + SecVeg + Wetland + Upland)
```

After the models have been selected we can see the selected models for each 
species, and find the predicted diversity for each site.
